---
title: "Traveler's Insurance"
author: "Tina, Jiayi, Jeremy"
format: html
editor: visual
---

For this case competition, your group's task is to provide a method for predicting the claim cost for each policy and to convince your business partner that your predictions will work well.

The InsNova data set is based on one-year vehicle insurance policies from 2004 to 2005. There are 45,239 policies, of which around 6.8% had at least one claim. The data is split to two parts: training data and validation data. In the validation and holdout data, `claimcst0` (claim cost), `clm` (claim indicator) and `numclaims` (claim counts) are omitted. You can build your model on the training data. In the end, use your best model to score the validation and holdout data. We will evaluate your model based on your validation and holdout data prediction.

Omitted Variable:

-   `clm` - Indicator of claim (0=no, 1=yes)

-   `numclaims` - The number of claims

-   `id` - Policy key

Outcome Variable:

`claimcst0` - Claim Cost

Predictor Variable:

-   `veh_value` - Market value of the vehicle in \$10,000's

-   `exposure` - The basic unit of risk underlying an insurance premium

-   `veh_body` - Type of vehicles

-   `veh_age` - Age of vehicles (1=youngest, 4=oldest)

-   `gender` - Gender of driver

-   `area` - Driving area of residence

-   `agecat` - Driver's age category from young (1) to old (6)Driver's age category from young (1) to old (6)

-   `engine_type` - Engine type of vehicles

-   `max_power` - Max horsepower of vehicles

-   `driving_history_score` - Driving score based on past driving history (higher the better)

-   `veh_color` - Color of vehicles

-   `marital_status` - Marital Status of driver (M = married, S = single)

-   `e_bill` - Indicator for paperless billing (0 = no, 1 = yes)

-   `time_of_week_driven` - Most frequent driving date of the week (weekdays vs weekend)

-   `time_driven` - Most frequent driving time of the day

-   `trm_len` - term length (6-month vs 12-month policies)

-   `credit_score` - Credit score

-   `high_education_ind` - indicator for higher education

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.show='hide'}
library(readr)
library(dplyr)
library(tidyverse)
library(geosphere)
library(car)
library(corrplot)
library(caret)
library(knitr)
library(sjPlot)
library("ISLR")
library("SmartEDA")
library(DataExplorer)
```

```{r}
nova <- read_csv("InsNova_data_2023_train.csv")
view(nova)
glimpse(nova)
```

```{r}
plot_bar(nova)
```

```{r}
plot_histogram(nova)
```

```{r}
# Structure of the data - Type = 1
ExpData(data=nova, type=1)

# Structure of the data - Type = 2
ExpData(data=nova,type=2)
```

```{r}
na_counts <- sapply(nova, function(x) sum(is.na(x)))
print(na_counts)
na_counts_with_na <- na_counts[na_counts > 0]
print(na_counts_with_na)
```

Total Policies: 22619

```{r}
summary(nova$id)
```

Market Value of the vehicles in \$10,000's

```{r}
summary(nova$veh_value)
```

```{r}
hist(nova$veh_value, main="Histogram of Market Value of the vehicles in $10,000's", xlab="Vehicle Value", col="lightblue", border="black")
```
